---
title: "Reporting With `R`"
subtitle: "It's easier than you think"
author: "Alex Axthelm"
date: "March 22, 2018"
output:
  xaringan::moon_reader:
    css: ["default", "default-fonts", "hygge"]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Why use R

- Reproducible

--

- Transformable

--

- Easier than you think

---
# About me

- Alex Axthelm
- Data Analyst at Indiana Commsission for Higher Education
- Masters Student in Data Science at IUB
- Works with `R` regularly  (~5 years experience)

---
layout: true

# Setting up your workspace

---
class: inverse
.center[.Large[Before we get started, Let's get ready.]]
---

## Opening RStudio

- RStudio Layout
- Navigating Multi-panes
- Source vs. Console

---
## RStudio Settings -> General 

I have strong opinons about best practices, based on experience.

You are free to ignore the settings I suggest here, but all of these are designed to make your life easier.

As we go along, there will also be opinons such as:

.full-width[.content-box-purple[Use indentation to show logical progression]]

---

From the menu bar: `Tools` -> `Global Options`

- **UNCHECK** Re-use idle sessions for project links
- **UNCHECK** Restore .RData into workspace at startup
  -- Save workspace to .RData on exit: **SET TO NEVER**

---
## RStudio Settings -> Code

### Editing
- **UNCHECK** Soft-wrap R source files

---
## RStudio Settings -> Code

### Display
- **CHECK** Show margin

.full-width[.content-box-purple[
Making your lines only 80 characters wide means you don't need to scroll to the side
]]

- **CHECK** Show indent guides

---
## `Lint`ing

> A linter or lint refers to tools that analyze source code to flag programming errors, bugs, stylistic errors, and suspicious constructs.
> 
> --https://en.wikipedia.org/wiki/Lint_(software)

We will use a package called `lintr` to help warn us of errors, and encourage good habits while coding.

To Enable `lintr`:

.content-box-green[`Tools` -> `Global Options` -> `Code` -> `Diagnostics`]

---
## Good practices to follow:

- Clearly defined, but succinct variable names

---
layout: false

# Our goal today

- Start with some "raw" data
  - (in Excel, no less)
- Transform it to extract useful information
- Make some nice plots
- Wrap everything into a human-readable report
- Keep our sanity throughout
---
layout: true

# R basics
---
class: inverse
.center[.Large[Let's explore some of the basics of `R`.]]
---
## Assigning Objects

`R` is an untyped language, with no initialization required.

```{r}
x <- 10
y <- 20
x + y
```
--
```{r}
z <- x + y
```
--
```{r}
z
print(z)
```
---
layout: true

# Importing

---
class: inverse
.center[.Large[Now we're really on our way.]]
---

---
# Our data set

Source Data for ICHE's "College Readiness Report" ([http://in.gov/che/2489.htm](http://in.gov/che/2489.htm))
